Containers and R: the Rockerverse and beyond

### 2019-10-07 manuscript.Rmd
part about r-ci

- [`rize` for Shiny](https://github.com/cole-brokamp/rize)

One of the key tool of Docker are dockerfiles, which can be thought of as "recipes" used to build a specific image. 
But building these files can feel like a workflow break as it demands to open a separate file, and / or to write a small wrapper around R's `write()` function to add elements to the file. 
At the same time, doing it by hand prevents from using programming language for what they are good for: iteration, and automation. 
Iteration and automation for Dockerfiles is the very reason behind [`dockerfiler`](https://github.com/ColinFay/dockerfiler), an R package designed for building Dockerfiles straigth from R.

For example, the [`golem`](https://github.com/ThinkR-open/golem) package (an R package for building production-grade Shiny application) makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for deploying the Shiny application to production. 
The reason behind this is that `dockerfiler`, on top of being scriptable from R, can leverage all the tools available in R to parse a `DESCRIPTION` file, to get system requirements, to list dependencies, versions, etc.

@@

- R-Hub
- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/
- `r-ci`: https://github.com/ColinFay/r-ci
- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.
- Google Cloud Build [@MarkEdmondson1234]
  - https://cloud.google.com/cloud-build/
  - For me this is what makes Docker containers viable, as it builds the Dockerfiles on each GitHub commit.  It couples with Google Container Registry to build private Docker images, for downstream applications.

>>>>

- [`rize` for Shiny](https://github.com/cole-brokamp/rize)
- [`dockerfiler`](https://github.com/ColinFay/dockerfiler)

@@

- R-Hub
- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/

When doing continuous integration and continuous delivery, it's crucial to test in an environment that matches the production environment.
Tools like `Gitlab-CI` are built on top of Docker images: the user specifies a base Docker image, and the whole tests are run inside this environment. 
But, as we just said, this environment has to be fixed, but it also have to contain the necessary toolkit. 
In order to achieve that, [`r-ci`](https://github.com/ColinFay/r-ci) combines `rocker` versioning and a series of tools specifically designed for testing. 
That way, package builders can use this image as a base image for there testing environment, without having to install the necessary packages every time they need to run a new test.

- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.
- Google Cloud Build [@MarkEdmondson1234]
  - https://cloud.google.com/cloud-build/
  - For me this is what makes Docker containers viable, as it builds the Dockerfiles on each GitHub commit.  It couples with Google Container Registry to build private Docker images, for downstream applications.
